Better Corpus Visualization through Topic Ranking

### 2011-02-10 topic_validation/topic_ranking_nips/paper.tex
Moved some files because I'm changing my thesis

\title{Better Corpus Visualization through Topic Ranking}
\author{
Matthew J. Gardner \\
Department of Computer Science \\
Brigham Young University \\
Provo, UT 84604 \\
\texttt{mjg82@byu.edu} \\
\And
Other authers\ldots \\
Affiliation \\
Address \\
\texttt{email} \\
}

%\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}
\maketitle

@@

Topic modeling is a popular and effective method for gathering meaning from a
set of documents.  Such models contain a lot of information and can be very
difficult to analyze.  For the proper visualization of these models, one must
present only the topics that are relevant to the task at hand.  We present the
novel task of topic ranking in response to user questions that greatly aids in
the visualization of topic models.  We make use of the document distribution of
a topic in order to rank the topics according to specific user questions.

@@

\section{Introdcution}
\label{sec:introduction}

Topic modeling has been shown to be an effective means of extracting semantic
meaning from a collection of
documents~\cite{griffiths-2004-finding-scientific-topics}.  Topic models have
been used as an inputs to other kinds of algorithms, such as classification and
sentiment tasks~\cite{blei-2003-latent-dirichlet-allocation,
brody-2010-aspect-sentiment-model-for-reviews}, but the usefulness of topic
models as visualization tools in and of themselves has not been well
established, and is even in
question~\cite{newman-2010-visualizing-with-topic-maps}.  Some recent work has
been done in an attempt to make corpus browsers using the results of LDA, and
to make the results of LDA more easily interpretable, but there is much work
yet to be done~\cite{blei-arxiv-corpus-browser,
newman-2010-visualizing-with-topic-maps,
mei-2007-automatic-labeling-of-topic-models}.

Common approaches to using topic models to facilitate corpus browsing and
visualization are limited mainly to showing the similarity between documents in
the corpus in various ways.  Such methods are useful in their own right when
one is individually scanning documents, but they are insufficient to answer
broader questions about the corpus, like, ``What topics show up most in these
documents?'' and, ``What does a particular author in the corpus talk about most
frequently that is unique to that author?''  Those questions turn one away from
analyzing individual documents to trying to get information out of the topics
themselves.

Upon looking at the topics one is confronted with tens or hundreds of lists of
words, sorted by a number given to each list by the algorithm.  Sifting through
those lists and trying to understand what they mean, and whether or not a
particular topic is important, can be a daunting task, particularly to one not
very familiar with the algorithm.  The common practice of cherry-picking good
topics to present in a paper hides the fact that a large number of topics
returned by topic models are very hard to interpret or even completely
meaningless to humans.

In order for topic models to be useful for corpus browsing, some sense has to
be made out of the topics.  One recently proposed method is to assign to each
topic a coherence value, so that the person can ignore topics that are likely
to be meaningless~\cite{newman-2010-automatic-evaluation-of-topic-coherence}.
While topic coherence is able to eliminate meaningless topics from
consideration, it is not able to address many broader questions that a person
browsing the corpus may have.

In this paper we present the novel task of ranking topics by relevance to
specific user questions.  If one is to successfully visualize a corpus in a way
that answers broad questions, and if topics, not documents, are the natural
answer to those questions, then one must properly rank topics such that the
most prominent topics contain the answer to the user's questions.  We recognize
that not all questions about a corpus fit into this visualization model, but
there are many that do, such as those listed above.

The rest of this paper is outlined as follows.  We first give further
motivation and a more detailed description of the task of ranking topics by
relevance in \secref{ranking}.  \secref{docdist} describes a method for
determining the relevance of a topic to various possible questions a person
could have when looking at a corpus, and \secref{examples} shows examples of
how this technique provides a proper ranking.  Finally, in \secref{conclusion}
we conclude.

@@

We assert that for many tasks, the proper visualization of a topic model is a
ranking of topics, not simply a window onto individual documents.  When one
wants to see what topics are prevelant in a corpus, the answer is not found by
scanning individual documents, but by looking at the topics produced by a topic
model.  Likewise, when one wants to see trends of topics over time, one does
not look through the documents, but at the topics, computing some statistics of
those topics for each year for which there are documents.  Thus a proper
visualization of the topic model for these tasks is to show most prominently
those topics that are most relevant to the task.

We can formalize the topic ranking problem simply as follows.  Given a set of
topics from a topic model of a corpus, and a user question about the corpus,
rank the topics by relevance to the user's question.  This is very similar to
basic information retrieval tasks, but also significantly different, in that we
are not looking for a ranking of documents as the result of a query, but a
ranking of topics which answer broader questions about the corpus as a whole.

Clearly the topic ranking task is fundamental to the proper visualization of
topic models, but as far as we are aware, this is the first attempt to even
consider the problem.  We propose a very simple way to answer some user
questions through the use of the document distribution of each topic.  This
technique is by no means a general solution to the problem, but for some
specific tasks it provides rankings that gives users the topics that answer
their questions.

@@

The document distribution for a topic can be thought of as follows: suppose we
have a token labeled with a given topic; what is the probability distribution
over documents in the corpus that this token could have come from?  We
approximate this distribution with a sample from a Gibbs sampler of LDA.  The
distribution is thus defined as
\begin{align}
  \label{eq:entropy}
  p(d|z) = \frac{C(d,z)}{\sum_{d \in \mathcal{D}} C(d,z)}
\end{align}
where $d$ is a particular document, $z$ is a particular topic, $C(d,z)$ is the
number of tokens in document $d$ labeled with topic $z$, and $\mathcal{D}$ is
the corpus.

@@

Also, one can compute the distribution over any subset of the documents, where
$p(d|z)$ is defined as in \eqref{eq:entropy}, but instead of the entire corpus,
$\mathcal{D}$ represents some specific subset.  When using subsets of the
documents, we only consider topics which have at least one token in the given
subset, so as to avoid division by zero.

@@

The first question we address is, What topics are most prevelant in the corpus?
This is a simple question with an obvious answer that does not require using
$p(d|z)$.  One can simply order the topics by the number of tokens in the
topic.  Upon doing so one finds the following topics (we show the topic number
and the top ten words in the topic for the top five topics):

\begin{tabular}{l|l}
  35 & years time made long today place home day ago back \\
  42 & people don ve make country ll time put didn money \\
  119 & make national work effort including protect adiministration clear
	support policy \\
  15 & america world american future challenges great lead generation nation
	leadership \\
  108 & president work day senator house make white clinton care support \\
\end{tabular}

@@

\begin{tabular}{l|l}
  42 & people don ve make country ll time put didn money \\
  119 & make national work effort including protect adiministration clear
	support policy \\
  120 & law government rights respect case society fact laws principles order \\
  143 & jobs ve create economy make years billion companies invest energy \\
  101 & children family day man mother care love father parents young \\
\end{tabular}

@@

\begin{tabular}{l|l}
  129 & mccain news john senator obama remarks america events donate email \\
  119 & make national work effort including protect adiministration clear
	support policy \\
  27 & mccain news john search media speeches www johnmccain email freedom \\
  90 & war iraq honor defeat enemy greater responsibility responsibilities
	difficult great \\
  66 & dollars reform federal washington congress reforms public policy
	problems system \\
\end{tabular}

@@

Displaying the right topics to a user is essential to the effective
visualization of topic models.  We have presented the novel task of ranking
topics according to relevance to certain user questions.  We have also given a
method for ranking topics in order to answer a few specific questions.  Such
rankings allow visualizations of topic models that present information to users
in much more effective ways than was previously possible.

While our methods are useful for a few user questions, it remains as future
work to devise a method for ranking topics in answer to arbitrary questions.
The evaluation of such rankings is also a difficult task that must be left to
future work.

>>>>

\title{Better Corpus Visualization through Topic Ranking}
\author{
Matthew Gardner \\
Department of Computer Science \\
Brigham Young University \\
\texttt{mjg82@byu.edu} \\
\And
Kevin Seppi \\
Department of Computer Science \\
Brigham Young University \\
\texttt{kseppi@cs.byu.edu} \\
\And
Eric Ringger \\
Department of Computer Science \\
Brigham Young University \\
\texttt{ringger@cs.byu.edu} \\
}

\begin{document}
\nipsfinalcopy % Uncomment for camera-ready version
\maketitle

@@

Topic modeling is a popular and effective method for gathering meaning from a
set of documents.  Such models contain a lot of information and can be very
difficult to analyze.  Current attempts at using topic models to enhance
information discovery in a document collection are limited mainly to simple
corpus browsers.  However, there are many questions a person may have about a
corpus that are best answered by the topics found by the topic model.  To get
these kinds of information from the model, one must decide which topics best
answer the question and present only the best topics.  In an attempt to address
this issue, we present the novel task of topic ranking in response to these
kinds of user questions and give some methods that are useful in this task.

@@

\section{Introduction}
\label{sec:introduction}

There are many kinds of questions a person might have when confronted with a
previously unseen collection of documents.  Some of those questions are
factual, such as, ``Who was Barack Obama's campaign advisor?''  Some questions
look for documents, such as ``What speeches did John McCain give about the war
in Iraq?''  Still other questions ask about topics and topical trends in the
corpus, such as, ``What were Obama's consistent campaign themes?''  Each kind
of question requires a separate kind of visualization.  The first two kinds of
questions listed require facts and documents that answer the question, and the
methods of information retrieval are well suited to those questions.  The last
kind of question, however, naturally falls into the domain of topic modeling.

Topic modeling has been shown to be an effective means of extracting semantic
meaning from a collection of
documents~\cite{griffiths-2004-finding-scientific-topics}.  Topic models have
been used as an inputs to other kinds of algorithms, such as classification and
sentiment tasks~\cite{blei-2003-latent-dirichlet-allocation,
brody-2010-aspect-sentiment-model-for-reviews}, but, as far as we are aware,
their use in revealing information about a corpus has previously been limited
to corpus browsing~\cite{newman-2010-visualizing-with-topic-maps,
blei-arxiv-corpus-browser}.  While topic-based corpus browsers give a better
window into the documents in the corpus and can help to visualize some kinds of
information, they are not capable of effectively answering questions about the
topics in the corpus.

In order to properly visualize a topic model in response to topical questions,
one must present information about the topics themselves, not the documents,
and decide which topics are most relevant to the information desired by the
user.  Typical methods of getting such information from a topic model require
the user to sift through tens or hundreds of lists of words, trying to
understand what they mean, and decide whether or not a particular topic is
relevant.  This can be a daunting task.  The common practice of hand-picking
good topics to present in a paper hides the fact that a large number of topics
returned by topic models are very hard to interpret or even completely
meaningless to humans.

Clearly, then, for topic models to be useful for visualizing corpora in
response to these kinds of questions, some sense has to be made out of the
topics.  One recently proposed method is to assign to each topic a coherence
value, so that the person can ignore topics that are likely to be
meaningless~\cite{newman-2010-automatic-evaluation-of-topic-coherence}.  While
topic coherence is able to eliminate meaningless topics from consideration, it
is not able to address many broader questions that a person browsing the corpus
may have.

In an attempt to address this issue, we present the novel task of topic
ranking; that is, deciding which topics are most pertinent to a user's
topically-based question.  In general this is a very difficult problem, but it
is essential to the proper visualization of topic models for many applications.
Surprisingly, it appears that researchers have not as yet attempted to solve
this problem.  In this paper we formalize the topic ranking problem and present
some initial attempts at solving a limited subset of possible questions.

The rest of this paper is outlined as follows.  We first give further
motivation and a more detailed description of the topic ranking task in
\secref{ranking}.  \secref{docdist} describes a method for determining the
relevance of a topic to a few types of topically-based questions a person could
have when looking at a corpus, and \secref{examples} shows examples of how this
technique provides a proper ranking.  Finally, in \secref{conclusion} we
conclude.

@@

We assert that for many tasks, the proper visualization of a topic model is a
ranking of topics, not simply a window onto individual documents.  When one
wants to see what topics are prevalent in a corpus, the answer is not found by
scanning individual documents, but by looking at the topics produced by a topic
model.  Likewise, when one wants to see trends of topics over time, one does
not look through the documents, but at the topics, computing some statistics of
those topics for each year for which there are documents.  Thus the correct
visualization of a topic model for these tasks includes showing most
prominently those topics that are most relevant to the task.

We can formalize the topic ranking problem simply as follows.  Given a set of
topics from a topic model of a corpus, and a topically-based question about the
corpus (that is, a question whose answer is naturally a topic, not a fact or a
document), rank the topics by relevance to the question.  This is similar to
some information retrieval tasks, but also significantly different, in that we
are not looking for a ranking of documents as the result of a query, but a
ranking of topics which answer broader questions about the corpus as a whole.

Clearly the topic ranking task is fundamental to the proper visualization of
topic models, but as far as we are aware, this is the first attempt to even
consider the problem.  We propose a very simple way to answer a few types of
topically-based questions through the use of the document distribution of each
topic.  This technique is by no means a general solution to the problem, but we
show it to be useful for visualizing at least a subset of the kinds of
information one might be interested in obtaining from a corpus.

@@

The document distribution for a topic can be thought of as follows: suppose we
have a token labeled with a given topic; what is the probability distribution
over documents in the corpus that this token could have come from?  We
approximate this distribution with a sample from a Gibbs sampler of LDA.  Our
approximation of this distribution is thus defined as
\begin{align}
  \label{eq:entropy}
  p(d|z) = \frac{C(d,z)}{\sum_{d \in \mathcal{D}} C(d,z)}
\end{align}
where $d$ is a particular document, $z$ is a particular topic, $C(d,z)$ is the
number of tokens in document $d$ labeled with topic $z$, and $\mathcal{D}$ is
the corpus.

@@

Also, one can compute the distribution over any subset of the documents, where
$p(d|z)$ is defined as in \eqref{eq:entropy}, but instead of the entire corpus,
$\mathcal{D}$ represents some specific subset (in our applications the subset
always corresponds to a set of documents having some particular attribute, such
as a common author).  When using subsets of the documents, we only consider
topics which have at least one token in the given subset, so as to avoid
division by zero.

@@

The first question we address is, What topics are most prevalent in the corpus?
This is a simple question with an obvious answer that does not require using
$p(d|z)$.  One can simply order the topics by the number of tokens in the
topic.  Upon doing so one finds the following topics (we show the topic number
and the top ten words in the topic for the top five topics):

\begin{tabular}{l|l}
  35 & years time made long today place home day ago back \\
  42 & people don ve make country ll time put didn money \\
  119 & make national work effort including protect administration clear
	support policy \\
  15 & america world american future challenges great lead generation nation
	leadership \\
  108 & president work day senator house make white clinton care support \\
\end{tabular}

@@

\begin{tabular}{l|l}
  42 & people don ve make country ll time put didn money \\
  119 & make national work effort including protect administration clear
	support policy \\
  120 & law government rights respect case society fact laws principles order \\
  143 & jobs ve create economy make years billion companies invest energy \\
  101 & children family day man mother care love father parents young \\
\end{tabular}

@@

\begin{tabular}{l|l}
  129 & mccain news john senator obama remarks america events donate email \\
  119 & make national work effort including protect administration clear
	support policy \\
  27 & mccain news john search media speeches www johnmccain email freedom \\
  90 & war iraq honor defeat enemy greater responsibility responsibilities
	difficult great \\
  66 & dollars reform federal washington congress reforms public policy
	problems system \\
\end{tabular}

@@

Many kinds of information a person might desire to get from a collection of
documents are best presented as topics from a topic model.  In order to
determine the relevance of each topic to specific user questions, we have
presented the novel task of topic ranking.  We have presented methods to answer
questions of the form, ``What are the consistent themes in this corpus, or for
this author?''  We have also shown results on a corpus of political speeches
that intuitively seem correct.

We believe that the topic ranking task is an important problem that deserves
more attention than it has as yet received.  This work is very preliminary in
that it only addresses a very small subset of possible topically-based
questions.  We plan on continuing to explore topic ranking, expanding the set
of questions that can be answered and searching for automated means of
evaluating these rankings.
